---
title: "p8105_HW3_sl5010"
author: "Fiona Li"
date: "10/16/2021"
output: github_document
---
## Problem 1

```{r}
library(dplyr)
library(tidyverse)
library(p8105.datasets)
data("instacart")
skimr::skim(instacart)
```

**size and structure of the data**: Number of rows is 1384617, Number of columns is 15. **key variables**: aisle, product_name, , order_dow, order_hour_of_day.

### How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>%
  group_by(aisle) %>%
  summarise(n_obs = n()) %>%
  mutate (ranking = min_rank(n_obs))
```
Hence, there are 134 aisles, "fresh vegetables" aisle is the most items ordered from, which have 	150609 items.

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>%
  group_by(aisle) %>%
  summarise(n_obs = n()) %>%
  filter(n_obs > 10000) %>%
  ggplot(aes(x = n_obs, y = aisle)) + geom_point() +
  labs(title = " the number of items ordered in each aisle")
```

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  group_by(aisle, product_name) %>%
  summarise(n_obs = n()) %>%
  filter(aisle == "baking ingredients"| aisle == "dog food care"| aisle == "packaged vegetables fruits") %>%
  top_n(1,n_obs) %>%
  pivot_wider(
    names_from = "product_name", 
    values_from = "n_obs"
  )
```
### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_order_hour_of_day = mean(order_hour_of_day)) %>%
  filter(product_name == "Pink Lady Apples"| product_name == "Coffee Ice Cream") %>%
  pivot_wider(
    names_from = "order_dow", 
    values_from = "mean_order_hour_of_day"
  )
```

## Problem 2

```{r}
library(p8105.datasets)
  data("brfss_smart2010")
```

### do some data cleaning:

```{r}
## format the data to use appropriate variable names
janitor::clean_names(brfss_smart2010)
## focus on the “Overall Health” topic
brfss_smart2010 = 
brfss_smart2010 %>%
  ## include only responses from “Excellent” to “Poor”
  filter(Topic == "Overall Health", Response == "Excellent"|Response =="Poor" |Response == "Fair"|Response == "Very good"|Response =="Good")
## organize responses as a factor taking levels ordered from “Poor” to “Excellent”
brfss_smart2010$Response = as.factor(brfss_smart2010$Response)
```

### In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_smart2010 %>%
  filter(Year == 2002) %>%
  select(Locationabbr, Locationdesc) %>%
  group_by(Locationabbr) %>%
  summarise(num_location = length(unique(Locationdesc))) %>%
  filter(num_location >= 7)
```
Hence, states were observed at 7 or more locations are : CT, FL, MA, NC, NJ, PA in 2002.

```{r}
brfss_smart2010 %>%
  filter(Year == 2010) %>%
  select(Locationabbr, Locationdesc) %>%
  group_by(Locationabbr) %>%
  summarise(num_location = length(unique(Locationdesc))) %>%
  filter(num_location >= 7)
```
Hence, states were observed at 7 or more locations are : CA, Co, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA in 2010.

### Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state

```{r}
brfss_smart2010 %>%
  filter(Response == "Excellent") %>%
  select(Year, Locationabbr, Data_value) %>%
  group_by(Year, Locationabbr) %>%
  summarise(avg_data_value = mean(Data_value)) %>%
  ggplot( aes(x = Year, y = avg_data_value, color = Locationabbr)) + 
  geom_line() +
  labs(title = " average data_value over time within a state")
```

### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_smart2010 %>%
  filter(Year == 2006 | Year == 2010, Locationabbr == "NY") %>%
  select(Year,Response, Data_value) %>%
  group_by(Response, Year) %>%
  summarise(avg_data_value = mean(Data_value)) %>%
  ggplot( aes(x = Response, y = avg_data_value, color = Year)) + geom_point() + facet_grid(. ~ Year) + labs(title = "two-panel plot for the years 2006, and 2010 of distribution of data_value for responses among locations in NY State")
```


```{r}

```

```{r}

```

```{r}

```
